Bayesian exponential random graph modeling of KWL Burial network

### 2020-05-30 analysis/paper/paper.Rmd
udpate background section

<!--archaeological setting-->
The burial data we analyzed are collected from the upper component of Kiwulan site, an Iron Age settlement site (AD1350-AD1950) located in northeastern Taiwan, which experienced the European colonial impacts in the 17th century and the large wave of Chinese immigrants in the 19th century. The excavation revealed abundant pottery sherds, imported ceramics and stonewares, wooden artifacts, stone tools, metal artifacts, imported glass beads and agates beads, and pipes [@Chen2007]. In addition to these artifacts, 90 burials, hundreds of middens, storage pits, and postholes with in-situ posts were also excavated. The burials were mostly located in the AD section, which is the largest open area at Kiwulan site that provides continuous stratigraphic sections suitable for temporal comparison. Those burials are oriented in an east-west direction located on the north side of the house structures indicated by postholes and `in-situ` wooden posts.

Previous studies on burials suggest a presence of uneven distribution of prestige goods across burials, accompanied with a debate over the differentiation represent an economic inequality supporting an egalitarian society or political inequality indicating a stratified society [@Cheng2008; @Hsieh2012]. @Cheng2008 interprets the unequal distribution of prestige materials, such as the golden beads, between burials, as evidence for hierarchy. However, @Hsieh2012 suggests a relatively egalitarian form from the comparative analysis of frequency and proportion of burial goods. She finds that the rich burials were usually associated with elders, which might indicate achieved, rather than inherited, status. The use of prestige goods in burials might be traced back before the European arrival according to historical records from Spanish and Chinese [@Borao2009; @LiandWu2006). Analysing burial data will give an insight into the social reactions of Kiwulan residents and provide a way to examine and explain the changes in social inequality at Kiwulan before and after European contact.

<!--chronology-->
Burials were primarily assigned to the pre-European period, European and post-European period, and Chinese period for network comparison. Since the sedimentary characteristics show vertical and lateral continuity cross units of AD section, depth is used as the first criterion for burial chronology. The chronology is then evaluated by the burials with light grey glazed jars, also called An-ping jars produced most likely in China and circulated in Southeast Asia from the late 16th to 18th century but are often found at sites with the European occupation in the early 17th century in Taiwan [@Berrocal2018; @Hsieh1995; @Ketel2011]. In addition to the diagnostic porcelain, three radiocarbon dates of charcoal associated with three different burials are used as another evidence to evaluate the chronology [@Chen2007]. Some burials that were heavily disturbed by modern construction are excluded to avoid misinterpreting from incomplete data.

<!--unanswered issues-->
The archaeological sites in northeastern Taiwan show some evidence of accumulation of imported prestige goods in burials [@Cheng2008], which might indicate the pursuit of prestige or wealth. In addition to prestige, European colonial powers could have provided military support to indigenous allies. For example, Voss’s (2005) study of a colonial outpost in California shows that colonial residents have different identities and higher social status than other local indigenous people. A general model to summarize these case studies might be that the support of a colonial power combined with high local values for imported goods, might lead to competition among individuals. In this study, I will use this model of competition to explain the changes in social inequality at Kiwulan before and after European contact.

@@

The network statistics presents that each θ corresponds to the parameter specified in ERGM previously. In general, positive mean indicates positive correlation, while negative mean indicates negative correlation.

>>>>

<!--archaeological setting-->
We analyzed burial data collected from the upper component of Kiwulan site, an Iron Age settlement site (AD1350-AD1950) located in northeastern Taiwan, which experienced the European colonial impacts in the early 17th century and the large wave of Chinese immigrants in the 19th century [@Chen2007]. The excavation revealed abundant pottery sherds, imported ceramics and stonewares, wooden artifacts, stone tools, metal artifacts, imported glass beads and agates beads, and pipes [@Chen2007]. In addition to these artifacts, 90 burials, hundreds of middens, storage pits, and postholes with in-situ posts were also excavated. The burials are mostly located in the AD section, which is the largest open area at Kiwulan that provides continuous stratigraphic sections suitable for temporal comparison. Those burials are oriented in an east-west direction on the north side of the house structures indicated by postholes and `in-situ` wooden posts, indicating a well organized spatial arrangement.

Previous studies on burials suggest a presence of uneven distribution of prestige goods across burials. However, there is a debate over whether the differentiation between burials hints an economic inequality supporting an egalitarian society [@Hsieh2012], or political inequality indicating a stratified society [@Cheng2008]. @Cheng2008 interprets the unequal distribution of prestige materials, especially the gold-foil beads between burials, as evidence for hierarchy. However, @Hsieh2012 suggests a relatively egalitarian society based on the comparative analysis of frequency and proportion of all burial goods. She found that the burials with rich goods were usually associated with elders, which might indicate achieved, rather than inherited, status. The use of prestige goods in burials might be traced back before the European arrival according to historical records from the Spanish and the Chinese [@Borao2009; @LiandWu2006). For example, the Spanish document by Spanish Priest documented that the Spanish soldier used agate beads to exchange materials from local indigenous people since they were valuable in their culture[Citation]. Also, there was a custom that people would bury imported materials with deceased to display their status and influence [Ciation]. Despite we observed the economic differentiation in burial goods and the value of imported goods according to historical documents, the extent of economic differentiation and how it changed over time remain unclear. Analyzing all burial data using network analysis approaches and compare the networks before and after the foreign presences will give an insight into the degree and changes in social inequality. This is important to understand the social reactions of Kiwulan residents in the context of foreign contact.

<!--chronology-->
We assigned burials to the pre-European period, European and post-European period for network comparison according to depth, radiocarbon dates, diagnostic materials, stratigraphic data, and previous studies. Since the sedimentary characteristics show vertical and lateral continuity cross units of AD section, depth is used as the first criterion for burial chronology. The chronology is then evaluated by the burials with time marker, light gray glazed jars, also called An-ping jars. The jars were believed to be produced in Southeast China and circulated in Southeast Asia from the late 16th to 18th century, but are often found at sites with the European occupation in the early 17th century in Taiwan [@Berrocal2018; @Hsieh1995; @Ketel2011]. In addition to the diagnostic jars, three radiocarbon dates of charcoal associated with three different burials are used as another evidence to evaluate the chronology [@Chen2007]. Some burials that were heavily disturbed by modern construction are excluded to avoid misinterpreting due to incomplete data.

<!--unanswered issues-->
The archaeological sites in northeastern Taiwan show some evidence of accumulation of imported prestige goods in burials around the same time as European contact [@Chen2007], which might indicate the pursuit of prestige or wealth in the society. In addition to prestige, European colonial powers could have provided military support to indigenous allies. For example, Voss’s (2005) study of a colonial outpost in California shows that colonial residents have different identities and higher social status than other local indigenous people. A general model to summarize these case studies might be that the support of a colonial power combined with high local values for imported goods, might lead to competition among individuals. In this study, We will use this model of competition to explain the changes in social inequality at Kiwulan before and after European contact.

@@

The network statistics presents each θ that corresponds to the parameter specified in ERGM previously. In general, positive mean indicates positive correlation, while negative mean indicates negative correlation.
